---
title: "BD N de Sant. GLM"
author: "Acevedo, et.al.,"
date: "5 de marzo de 2018"
output: html_document
df_print: paged
editor_options:
  chunk_output_type: console
---
## BD 
## Read packages
```{r setup, include=TRUE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sp)
library(rgdal)
library(raster)
library(rgdal)
library(dismo)
library(rasterVis)
library(xtable)
library(sf)
library(mapview)
library(corrgram)
library (corrplot) # correlations
library(lme4)
library(nlme)
# library(arm)
library(sjPlot)

```


### Read Data

```{r data1}
# Path to working directory
path <- "C:/Users/diego.lizcano/Documents/GitHub/Bd_N_Santander/data"
shp <- "C:/Users/diego.lizcano/Documents/GitHub/Bd_N_Santander/Data2"
bio <- "C:/Users/diego.lizcano/Documents/GitHub/Bd_N_Santander/Capas2"

BD4<-readOGR(dsn=shp, layer="Bd_generos1")

# define geo coords and put to SpatialPointDataFrames
crs.geo <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")  # geographical, datum WGS84
proj4string(BD4) <- crs.geo  # define projection system of our data

BD <- st_as_sf(BD4) # Convert foreign object to an sf object

pos <-  grep(" ", BD$Species)
keep <-  substr(pos, 1, pos-1)

########################
# BD$genus <- 

###cargando variables ambientales
files <- list.files(path=bio, pattern='tif', full.names=TRUE )

```


### View files

```{r data2}
# View(files)
# Agrupar las variables ambientales en un objeto stack
predictors <- stack(files)
### clac  slope aspect
slope_aspect <- terrain(predictors$DEM30, opt = c("slope", "aspect"), unit = "degrees")

predictors <- addLayer(predictors, slope_aspect)

deptos <-raster::getData ('GADM', country='COL', level=1)

# names(predictors)
# plot(predictors$Bio1)
# plot(BD1, pch=18, add=TRUE, col="blue", main="Batrachochytrium dendrobatidis")

mapview(BD, zcol = "Diagnostic", legend = TRUE, map.types = c("Esri.WorldShadedRelief", "Esri.WorldImagery"), color = "grey40")

# plot(BD["Diagnostic"], add=T)
# levelplot(predictors)

### correlation 
jnk=layerStats(predictors, 'pearson', na.rm=T)
corr_matrix=jnk$'pearson correlation coefficient'


# corrgram(corr_matrix, order=NULL, lower.panel=panel.shade,
#  upper.panel=NULL, text.panel=panel.txt,
#  main="Car")

##### correlation plot
# corrplot(corr_matrix, order = "FPC",type = "lower",  tl.cex = 0.8) # tl.pos = "lower",

####### function to add significance 
# mat : is a matrix of data
# ... : further arguments to pass to the native R cor.test function
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}

# matrix of the p-value of the correlation
# use the function
p.mat <- cor.mtest(predictors)
# head(p.mat[, 1:5])

##### correlation plot 
# Leave blank on no significant coefficient
corrplot(corr_matrix, p.mat = p.mat, sig.level = 0.05,  order="FPC", type = "lower",  tl.cex = 0.8) # tl.pos = "lower",


```


### Remove correlated layers

```{r data3}
## remove correlated layesr
predictors2 <- dropLayer(predictors, c(1,2,3,15,18,19))

### correlation 
jnk=layerStats(predictors2, 'pearson', na.rm=T)
corr_matrix=jnk$'pearson correlation coefficient'


# matrix of the p-value of the correlation
p.mat <- cor.mtest(predictors2)

corrplot(corr_matrix, p.mat = p.mat, sig.level = 0.05,  order="FPC", type = "lower",  tl.cex = 0.8) # tl.pos = "lower",

print("predictores no correlacionados")
names(predictors2)

```

### Building GLM models

```{r data4}

#Extraer la informacion de variables ambientales por ptos. con presencia de BD

# # head(BD)
# BD <- BD1[,-1] # quito la primera columna
# head(BD)
# names(BD)[1]<-paste("x")
# names(BD)[2]<-paste("y") # re nombro encabezados de  columnas
# # head(BD)
# 
# presvals <- (BD)
positive <- BD %>% dplyr::filter (Diagnostic==1) # %>% as.data.frame()
negative <- BD %>% dplyr::filter (Diagnostic==0) # %>% as.data.frame()

presvals <- as.data.frame(extract(scale(predictors2, center=TRUE, scale=TRUE), 
                    as(positive, "Spatial")))

ausvals <- as.data.frame(extract(scale(predictors2, center=TRUE, scale=TRUE), 
                   as(negative, "Spatial")))

###### genus adding 
presvals$genus <- positive$Genero
presvals$pa <-  rep(1,88)
ausvals$genus  <- negative$Genero
ausvals$pa <-  rep(0,62)
                 
########### Full data
fulldat <- rbind(presvals, ausvals) #####


# 
set.seed(2000)
# backg <- randomPoints(predictors, n=500, extf = 1.25)


# plot(predictors2$DEM30)
# plot(BD["Diagnostic"], add=T)
# plot(BD, pch=18, add=TRUE, col="blue", main="Batrachochytrium dendrobatidis")
# 

#generando datos de prueba y entrenamiento con los datos de presencia y pseudo-ausencia con particion 80-20%
indexes_training = sample(1:nrow(fulldat), size=0.70*nrow(fulldat))
indexes_testing = sample(1:nrow(fulldat), size=0.30*nrow(fulldat))

###Presencias
#identifica el 80% de los datos como datos training
training = as.data.frame(fulldat[indexes_training,])
dim(training)
#identifica el 20% de los datos como datos testing
testing = as.data.frame(fulldat[-indexes_training,])
dim(testing)

# species
# testing$species <- positive$Species[-indexes_pres]
# training$species <- positive$Species[indexes_pres]
# View(testing)


###ausencia
#identifica el 20% de los datos como datos prueba/test
backg_test = as.data.frame(ausvals[indexes_aus_test,])
dim(backg_test)
#identifica el 80% de los datos como datos entrenamiento/train
backg_train = as.data.frame(ausvals[-indexes_aus_test,])
dim(backg_train)

# species
# View(testing)
backg_train$species <- negative$Species[-indexes_aus_test]
backg_test$species <- negative$Species[indexes_aus_test]


# mapeando los datos de entrenamiento y prueba de presencias y ausvals
# r = raster(predictors2, 1)
# plot(!is.na(r), col=c('white', 'light grey'), legend=FALSE, main="datos de entrenamiento y prueba ")
# points(backg_train, pch='-', cex=0.5, col='red')
# points(backg_test, pch='-', cex=0.5, col='black')
# points(testing, pch= '+', col='green')
# points(training, pch='+', col='blue')
# 

trainpres <- as.data.frame(testing)# data.frame( extract(predictors2, testing) )
trainbackg <- as.data.frame(backg_train)# data.frame( extract(predictors2, backg_train) )
train <- rbind(trainpres, trainbackg)
pb_train <- c(rep(1, nrow(trainpres)), rep(0, nrow(trainbackg)))
envtrain <- data.frame( cbind(pa=pb_train, train) )


#### extract from stack
testpres <- as.data.frame(training)# data.frame( extract(predictors2, training) )
testbackg <- as.data.frame(backg_test)# data.frame( extract(predictors2, backg_test) )

# head(envtrain)


#Ajuste del modelo GLM y predicc??n 
# GLM binomial 
glm1 <- glm(pa ~ Bio12 +  Bio13 + Bio14 + Bio15 +
            Bio16 + Bio17 + Bio18 + Bio19 + Bio2 +  
            Bio3 + Bio4 + Bio6 + Bio7 + 
            DEM30 + slope + aspect,
           family = binomial(link = "logit"), data=training)
summary(glm1)
coef(glm1)


# Stepwise Regression
library(MASS)
step <- stepAIC(glm1, direction="backward", k = 2)
step$anova # display results


##########################################
## now only covars from Final Model glm1
##########################################

glm2 <- glm(pa ~ Bio12 + Bio13 + Bio14 + Bio17 + Bio19 + Bio2 + Bio4 + Bio6,
            family = binomial(link = "logit"), data=training)


glm3 <- glm(pa ~ Bio12 + Bio13 + Bio14 + Bio17 + Bio2 + Bio6,
            family = binomial(link = "logit"), data=training)

glm4 <- glm(pa ~ Bio12 + Bio13 + Bio19 + Bio2 + Bio6,
            family = binomial(link = "logit"), data=training)

glm5 <- glm(pa ~ Bio12 * Bio6 + Bio2 + Bio17,
            family = binomial(link = "logit"), data=training)

AIC(glm2, glm3, glm4, glm5)


summary(glm2)
#### multiple group effects with multiple group effect terms factor species.
mod_lmer1 <- glmer(pa ~ Bio12 + Bio13 + Bio14 + Bio17 + Bio2 + Bio6 + (1|genus), data=training, #fulldat,
                   family = binomial(link = "logit"))

################ plot using sjPlot
plot_model(mod_lmer1, vline.color = "gray", 
                      sort.est = TRUE)

sjp.lmer(mod_lmer1, type = "re")
# sjp.lmer(mod_lmer1, type = "slope")

## evaluacion
ge2 <- evaluate(testing[testing$pa==1,], testing[testing$pa==0,],  glm2)
ge2


ge_lmer <- evaluate(testing[testing$pa==1,], testing[testing$pa==0,], mod_lmer1, allow.new.levels = TRUE) ####### change by genus
ge_lmer

ge3 <- evaluate(testing[testing$pa==1,], testing[testing$pa==0,],  glm3, allow.new.levels = TRUE) ####### change by genus
ge3

par(mfrow=c(1,3))
plot(ge2, 'ROC', main="GLM")
plot(ge_lmer, 'ROC', main="GLMM")
plot(ge3, 'ROC', main="GLMM")
# dev.off() 






##mapeo del modelo GLM3
pg3 <- predict(scale(predictors2, center=TRUE, scale=TRUE), glm3, ext=predictors2)
par(mfrow=c(1,2))
plot(pg3, main='GLM3/binomial, BD')
plot(BD4, add=TRUE, col='grey50')
plot(deptos, add=TRUE, border='white')
tr <- threshold(ge3, 'spec_sens')
plot(pg3>tr, main='presence/absence')
plot(BD4, add=TRUE, col='grey50')
plot(deptos, add=TRUE, border='white')
# points(testing, pch='+')
# dev.off()

##mapeo del modelo GLM3
pg_lmer <- predict(scale(predictors2, center=TRUE, scale=TRUE), mod_lmer1, ext=predictors2)
# par(mfrow=c(1,2))
plot(pg_lmer, main='GLM/binomial, BD')
plot(BD4, add=TRUE, border='dark grey')
tr <- threshold(ge_lmer, 'spec_sens')
plot(pg>tr, main='presence/backgence')
plot(BD4, add=TRUE, border='dark grey')
points(testing, pch='+')

dev.off()



```

# Comentarios
BIO1 = Annual Mean Temperature
BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp))
BIO3 = Isothermality (BIO2/BIO7) (* 100)
BIO4 = Temperature Seasonality (standard deviation *100)
BIO5 = Max Temperature of Warmest Month
BIO6 = Min Temperature of Coldest Month
BIO7 = Temperature Annual Range (BIO5-BIO6)
BIO8 = Mean Temperature of Wettest Quarter
BIO9 = Mean Temperature of Driest Quarter
BIO10 = Mean Temperature of Warmest Quarter
BIO11 = Mean Temperature of Coldest Quarter
BIO12 = Annual Precipitation
BIO13 = Precipitation of Wettest Month
BIO14 = Precipitation of Driest Month
BIO15 = Precipitation Seasonality (Coefficient of Variation)
BIO16 = Precipitation of Wettest Quarter
BIO17 = Precipitation of Driest Quarter
BIO18 = Precipitation of Warmest Quarter
BIO19 = Precipitation of Coldest Quarter



